Bicycle Balance Assistance Reduces Probability of Falling at Low Speeds When Subjected to Mechanical Perturbations

### 2024-06-27 main.tex
b is not balance variable.

\section{Methods}
%
\subsection{Bicycle}
%
We modified a standard electric bicycle (Gazelle Grenoble C7\todo{Double check
this is the right model}, Dieren, The Netherlands) with a custom motor in the
steering assembly capable of applying up to 7~\si{\newton\meter} of torque
between the head tube and steer tube, see
Figure~\ref{fig:balance-assist-bicycle}.  A custom motor controller converts
the commanded torque to applied torque. We measure the speed of the rear wheel
with an encoder (Magura XXX) and measure the body fixed roll rate of the
bicycle with a MEMs rate gyroscope (CRS 03-02s, Silicon Sensing, United
Kingdom).  \todo{Do you mean the gyro inside the bicycle or the external?
Currently filled in the data of the external one. Not sure what make and model
is inside the bicycle} The balance assist control algorithm is implemented on a
microprocessor (Teensy, PJRC, USA) and data from all sensors is logged with a
CAN bus (CanEdge2, CSS Electronics, Denmark) at approximately 100 Hz.
\todo{Sample rate is not the same for each sensor: wheel speed, imu = 1ms,
steer torque command is updated every 10ms, can log is written every 10ms}

@@

\subsection{Statistics}
%
We test the hypothesis that the balance assistance controller will reduce the
probability of falling when perturbed externally at the handlebar. We have a
single binary fall outcome variable \(f\) that is dependent on several possible
explanatory independent variables, one of which is the binary balance
assistance (on or off). See Table~\ref{tab:stat-model-variables} for all of the
variables.
%
\begin{table}
  \centering
  \caption{Statistical model variables.}
  \label{tab:stat-model-variables}
  \begin{tabular}{lll}
    \toprule
    Variable & Units & Description \\
    \midrule
    \(L\) & \si{\newton\meter\second} & angular impulse of perturbation torque \\
    \(c\) & integer & order number of perturbation \\
    \(k\) & TODO & gain of balance assist control \\
    \(\delta_0\) & \si{\degree} & steer angle at start of perturbation \\
    \(\phi_0\) & \si{\degree} & roll angle at start of perturbation \\
    \(v\) & \si{\meter\per\second} & forward speed (1.7 or 2.8) \\
    \(f\) & boolean & outcome of perturbation: did not fall, did fall \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\section{Discussion}
%
The probability that a fall occurs depends on the values of all the independent
variables in Table~\ref{tab:stat-model-variables}, but we can visualize the
effect of one or two variables (e.g. Figure~\ref{fig:probability}) to gain
some insight.
But to interpret the results in Tables~\ref{tab:freq-coeffs-6} and
\ref{tab:freq-coeffs-10} it is important to understand the relationship between
probability and odds.
The estimate in Table~\ref{tab-freq-coeffs-6} shows that the balance assist
system halves the odds that a perturbation results in a fall
(MCiO=0.53)~\todo{Make a variable for MCiO}.
This means if the odds are a 1000:1, turning on the balance-assist system
reduces the odds to 500:1.
However, the probability that a fall occurs is only reduced from $0.999$ to
$0.998$ in that case.
If the odds that a fall will occur are smaller, halving the odds has a larger
influence on the fall probability.
For example, if the odds that a fall occurs is two, halving it to one reduces
the probability from $0.66$ to $0.50$.
This can be seen in Figure~\ref{fig:probablity} which shows the fall
probability as a function of the magnitude of the normalized angular impulse
for both speeds.
The difference in probability between the balance-assist on/off case becomes
insignificant outside of approximately \(\pm1\) standard deviation of the
average angular impulse.
This means that the balance assist is most effective for perturbations close to
the subject's personal threshold between falling or recovery and that large
perturbations will make you fall regardless of the balance assist's help.

>>>>

\section{Methods}
%
\subsection{Bicycle}
%
We modified a standard electric bicycle (Gazelle Grenoble C7\todo{Double check
this is the right model}, Dieren, The Netherlands) with a custom motor in the
steering assembly capable of applying up to 7~\si{\newton\meter} of torque
between the head tube and steer tube, see
Figure~\ref{fig:balance-assist-bicycle}.  A custom motor controller converts
the commanded torque to applied torque. We measure the speed of the rear wheel
with an encoder (Magura XXX) and measure the body fixed roll rate of the
bicycle with a MEMs rate gyroscope (CRS 03-02s, Silicon Sensing, United
Kingdom).  \todo{Do you mean the gyro inside the bicycle or the external?
Currently filled in the data of the external one. Not sure what make and model
is inside the bicycle} The balance assist control algorithm is implemented on a
microprocessor (Teensy, PJRC, USA) and data from all sensors is logged with a
CAN bus (CanEdge2, CSS Electronics, Denmark) at approximately 100 Hz.

@@

\subsection{Statistics}
%
We test the hypothesis that the balance assistance controller will reduce the
probability of falling when perturbed externally at the handlebar. We have a
single binary fall outcome variable \(f\) that is dependent on several possible
explanatory independent variables, one of which is the binary balance
assistance (on or off). See Table~\ref{tab:stat-model-variables} for all of the
variables.
%
\begin{table}
  \centering
  \caption{Statistical model variables.}
  \label{tab:stat-model-variables}
  \begin{tabular}{lll}
    \toprule
    Variable & Units & Description \\
    \midrule
    \(L\) & \si{\newton\meter\second} & angular impulse of perturbation torque \\
    \(c\) & integer & order number of perturbation \\
    \(\delta_0\) & \si{\degree} & steer angle at start of perturbation \\
    \(\phi_0\) & \si{\degree} & roll angle at start of perturbation \\
    \(v\) & \si{\meter\per\second} & forward speed (1.7 or 2.8) \\
    \(b\) & boolean & balance assist control on or off \\
    \(f\) & boolean & outcome of perturbation: did not fall, did fall \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\section{Discussion}
%
The probability that a fall occurs depends on the values of all the independent
variables in Table~\ref{tab:stat-model-variables}, but we can visualize the
effect of one or two variables (e.g. Figure~\ref{fig:probability}) to gain
some insight.
To interpret the results in Tables~\ref{tab:freq-coeffs-6} and
\ref{tab:freq-coeffs-10} it is important to understand the relationship between
probability and odds.
The estimate in Table~\ref{tab-freq-coeffs-6} shows that the balance assist
system halves the odds that a perturbation results in a fall
(MCiO=0.53)~\todo{Make a variable for MCiO}.
This means if the odds are a 1000:1, turning on the balance-assist system
reduces the odds to 500:1.
However, the probability that a fall occurs is only reduced from $0.999$ to
$0.998$ in that case.
If the odds that a fall will occur are smaller, halving the odds has a larger
influence on the fall probability.
For example, if the odds that a fall occurs is two, halving it to one reduces
the probability from $0.66$ to $0.50$.
This can be seen in Figure~\ref{fig:probablity} which shows the fall
probability as a function of the magnitude of the normalized angular impulse
for both speeds.
The difference in probability between the balance-assist on/off case becomes
insignificant outside of approximately \(\pm1\) standard deviation of the
average angular impulse.
This means that the balance assist is most effective for perturbations close to
the subject's personal threshold between falling or recovery and that large
perturbations will make you fall regardless of the balance assist's help.
